---
title: "Data 608 - Assignment 2"
author: "Habib Khan"
date: "February 23, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(ggplot2)
library(dplyr)
library(ggthemes)
library(kableExtra)
library(knitr)
library(dplyr)
library(grid)
library(gridExtra)
```


# Overview of Assignment

For module 2 we'll be looking at techniques for dealing with big data. In particular binning strategies and the datashader library (which possibly proves we'll never need to bin large data for visualization ever again.)
To demonstrate these concepts we'll be looking at the PLUTO dataset put out by New York City's department of city planning. PLUTO contains data about every tax lot in New York City.
PLUTO data can be downloaded from here. Unzip them to the same directory as this notebook, and you should be able to read them in using this (or very similar) code. Also take note of the data dictionary, it'll come in handy for this assignment.

# Getting and loading the data

```{r, eval=FALSE}
# url & file
base_url <- 'http://www1.nyc.gov/assets/planning/download/zip/data-maps/open-data/'
zip_file <- 'nyc_pluto_16v2%20.zip'

# download data 
download.file(paste0(base_url,zip_file),zip_file)

# unzip files
unzip(zip_file)

# subdirectory where unzipped file is stored
sub_dir <- 'BORO_zip_files_csv/'

# filename prefixes, each abbreviated to represent nyc boro
file_prefix <- c('BX','BK','MN','QN','SI')

# read csv data for each boro  
for (i in 1:length(file_prefix)) {
  
  # construct string: filepath, filename
  file_name <- paste0(sub_dir,file_prefix[i],".csv")
  
  # save each csv file as separate data frame
  assign(file_prefix[i],read.csv(file_name, stringsAsFactors = FALSE))
}

# save to master dataframe 
mydf <- rbind(BX, BK, MN, QN, SI)

# write master csv
write.csv(mydf, file="all_PLUTO_data.csv")
```

# Loading data from local directory

```{r}
data <- read.csv("all_PLUTO_data.csv", stringsAsFactors = FALSE)

kable(head(data)) %>% kable_styling()
```

